Single Bivalent Level Analysis

### 2020-01-24 doc/Bivalent_Analysis.Rmd
too much work on analysis

#1. how many biv's are from cells with >15 bivalents

over_15_biv_cell_DF <- cell.level.BivData_table[cell.level.BivData_table$Nbiv >15, ]
BivData_from_15biv_cells <- Curated_BivData[ Curated_BivData$fileName %in% over_15_biv_cell_DF$fileName, ]

@@

@@

mouse.avs_4MM <- ddply(Curated_BivData, c("mouse"), summarise,
                        Nmice = length(unique(mouse)),
                        Ncells  = length(Obj.ID),
                        mean_SC = as.numeric(format(round(  mean(chromosomeLength), 3 ), nsmall=3) ),
                        cV = cv(chromosomeLength),
                        var = as.numeric(format(round(   var(chromosomeLength),3), nsmall=3) ),
                        sd   = round(sd(chromosomeLength), 3),
                        se   = round(sd / sqrt(Ncells), 3),
                       #IFD things
                #  mean_IFD_ABS = as.numeric(format(round(  mean(IFD1_ABS, na.rm = TRUE), 3 ), nsmall=3) ),
                # mean_IFD_PER = as.numeric(format(round(  mean(IFD1_PER, na.rm = TRUE), 3 ), nsmall=3) ),

@@

@@

#vari
mouse.avs_4MM <- add_sex(mouse.avs_4MM)
mouse.avs_4MM <- add_strain(mouse.avs_4MM)
mouse.avs_4MM <- add_subsp(mouse.avs_4MM)
#unordered
mouse.avs_4MM$subsp <- factor(mouse.avs_4MM$subsp, ordered = FALSE)
mouse.avs_4MM$sex <- factor(mouse.avs_4MM$sex, ordered = FALSE)
mouse.avs_4MM$strain <- factor(mouse.avs_4MM$strain, ordered = FALSE)

#MAKE the same version of the TABLE but for the longest_biv_Dataset (when I get around to making it)
#code if males are high or low?
```

@@

3) Interference / IFD will not be sexually dimorphic. Previous measures has been shown to lack sexual dimorphism (cite petkov).

@@

```{r MM_2_SC.length, echo=FALSE, eval=FALSE}
#Long_biv_mouse.avs_4MM
#as.data.frame( round(  summary(BivData.MixedModel.HQ)$tTable, 6 ) )

@@

```{r log.reg_SC.length, echo=FALSE}

mouse.avs_4MM_male <- mouse.avs_4MM[mouse.avs_4MM$sex == "male",]
mouse.avs_4MM_M.male <- mouse.avs_4MM_male[mouse.avs_4MM_male$subsp == "Musc",]

#divide musc males into groups
mouse.avs_4MM_male$Rec.group <- ifelse(grepl("PWD", mouse.avs_4MM_male$strain), 1, 
                                    ifelse(grepl("MSM", mouse.avs_4MM_male$strain), 1, 0))
#

log.reg_SC.length <- glm(Rec.group ~ mean_SC, 
                            data=mouse.avs_4MM_male, family=binomial(link="logit"))

log.reg_SC.length_Musc <- glm(Rec.group ~ mean_SC, 
                            data=mouse.avs_4MM_male[mouse.avs_4MM_male$subsp == "Musc",], family=binomial(link="logit"))
mouse.av_log.reg_SC.length <- summary(log.reg_SC.length_Musc)

# bivalent level log regression
log.reg_SC.length_Musc_BIV <- glm(Rec.group ~ chromosomeLength, 
                            data=Curated_BivData[(Curated_BivData$subsp == "Musc")& (Curated_BivData$sex == "male"),], family=binomial(link="logit"))

bivalent_log.reg_SC.length <- summary(log.reg_SC.length_Musc_BIV)

@@

`r mouse.av_log.reg_SC.length$coefficients`

@@

`r bivalent_log.reg_SC.length$coefficients`

#### t.test for sc length

```{r t.test, echo=FALSE}

T.test_Poly_SC.Len_mouse.av <- t.test(mouse.avs_4MM_male$mean_SC[mouse.avs_4MM_male$Rec.group == 1],

  mouse.avs_4MM_male$mean_SC[(mouse.avs_4MM_male$subsp == "Musc")&(mouse.avs_4MM_male$sex == "male")&(mouse.avs_4MM_male$Rec.group == 0)]
)#p-value = 0.001567

T.test_Poly_SC.Len_single.biv <- t.test(Curated_BivData$chromosomeLength[Curated_BivData$Rec.group == 1],

  Curated_BivData$chromosomeLength[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male")&(Curated_BivData$Rec.group == 0)]
)#p-value < 2.2e-16

The mouse average and single bivalents are significantly longer in the high rec group p = `r T.test_Poly_SC.Len_mouse.av$p.value` and p=`r T.test_Poly_SC.Len_single.biv$p.value`.

@@

#isolate data
mouse.avs_4MM_male <- mouse.avs_4MM[mouse.avs_4MM$sex == "male",]
mouse.avs_4MM_M.male <- mouse.avs_4MM_male[mouse.avs_4MM_male$subsp == "Musc",]

@@

rand_SC.Length_MM.male.poly <- exactRLRT(lmer_MM_SC_length_male.poly)#this won't work either...hmm
```

@@

```{r show.IFD, echo=FALSE, warning=FALSE}

@@

```{r IFD.poly_log.reg, echo=FALSE}

#test. for the high vs low strains and IFD
#mouse.avs_4MM_male

@@

#Single Biv Leve;
single.biv.1CO_pos_t.test <- t.test(

  Curated_BivData$PER_Foci_1[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male")&(Curated_BivData$hand.foci.count == 1) & (Curated_BivData$Rec.group == 1)],

   Curated_BivData$PER_Foci_1[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male")&(Curated_BivData$hand.foci.count == 1) & (Curated_BivData$Rec.group == 0)]
  )

#0single biv level
```

@@

#Single Biv Level
t.test_single.biv_siscoten <- t.test(

@@

   Curated_BivData$SisCoTen[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male")& (Curated_BivData$Rec.group == 0)]
  )
```

>>>>

#1. how many biv's are from cells with >15 bivalents
over_15_biv_cell_DF <- cell.level.BivData_table[cell.level.BivData_table$Nbiv >15, ]
BivData_from_15biv_cells <- Curated_BivData[ Curated_BivData$fileName %in% over_15_biv_cell_DF$fileName, ]

@@

female.whole.cell <- whole.cell.data_manual[whole.cell.data_manual$sex == "female",]
data.frame(table(droplevels(female.whole.cell$category)))

#Here is the data set of female whole cell measures, apply rank to these measures ... 
#run a ddplyr to 
#

@@

mouse.avs_4MM <- ddply(Curated_BivData, c("mouse"), summarise,
                        Nmice = length(unique(mouse)),
                        Ncells  = length(Obj.ID),
                        mean_SC = as.numeric(format(round(  mean(chromosomeLength), 3 ), nsmall=3) ),
                        cV = cv(chromosomeLength),
                        var = as.numeric(format(round(   var(chromosomeLength),3), nsmall=3) ),
                        sd   = round(sd(chromosomeLength), 3),
                        se   = round(sd / sqrt(Ncells), 3),

                       #calq mean.sc by chrm class
                       mean.SC_1CO = mean(chromosomeLength[which(hand.foci.count == 1)], na.rm = TRUE),
                       mean.SC_2CO = mean(chromosomeLength[which(hand.foci.count == 2)], na.rm = TRUE),
                       mean.SC_3CO = mean(chromosomeLength[which(hand.foci.count == 3)], na.rm = TRUE),
                       mean.SC_4CO = mean(chromosomeLength[which(hand.foci.count == 4)], na.rm = TRUE),

                # IFD things
                #  mean_IFD_ABS = as.numeric(format(round(  mean(IFD1_ABS, na.rm = TRUE), 3 ), nsmall=3) ),
                # mean_IFD_PER = as.numeric(format(round(  mean(IFD1_PER, na.rm = TRUE), 3 ), nsmall=3) ),

@@

              mean.siscoten_1CO = mean(SisCoTen[which(hand.foci.count == 1)], na.rm = TRUE),
              mean.siscoten_2CO = mean(SisCoTen[which(hand.foci.count == 2)], na.rm = TRUE),
              mean.siscoten_3CO = mean(SisCoTen[which(hand.foci.count == 3)], na.rm = TRUE),

@@

#vari
mouse.avs_4MM <- add_sex(mouse.avs_4MM)
mouse.avs_4MM <- add_strain(mouse.avs_4MM)
mouse.avs_4MM <- add_subsp(mouse.avs_4MM)
mouse.avs_4MM <- add_category(mouse.avs_4MM)

#unordered for the mixed models (move this to the MM blocks...?)
#keep order for the plots?
mouse.avs_4MM$subsp <- factor(mouse.avs_4MM$subsp, ordered = FALSE)
mouse.avs_4MM$sex <- factor(mouse.avs_4MM$sex, ordered = FALSE)
mouse.avs_4MM$strain <- factor(mouse.avs_4MM$strain, ordered = FALSE)

#MAKE the same version of the TABLE but for the longest_biv_Dataset (when I get around to making it)
#code if males are high or low?

#assign high and low
#divide musc males into groups
mouse.avs_4MM$Rec.group <- ifelse(grepl("PWD male", mouse.avs_4MM$category), 1, 
                                    ifelse(grepl("MSM male", mouse.avs_4MM$category), 1, 0))

@@

3) Interference / IFD will not be sexually dimorphic. Previous physical measures of interference were not different between sexes (verify petkov 2001).

@@

```{r MM_2_SC.length_2_DELETE, echo=FALSE, eval=FALSE}
#Long_biv_mouse.avs_4MM
#as.data.frame( round(  summary(BivData.MixedModel.HQ)$tTable, 6 ) )

@@

```{r log.reg_SC.length, echo=FALSE}
#All chrm
log.reg_SC.length_mouse.av <- glm(Rec.group ~ mean_SC, 
                            data=mouse.avs_4MM[(mouse.avs_4MM$sex == "male") & (mouse.avs_4MM$subsp == "Musc"),]

                         , family=binomial(link="logit"))
summary(log.reg_SC.length_mouse.av)

#considering seperating out by hand foci counts

#1CO
log.reg_SC.length_mouse.av_1CO <- glm(Rec.group ~ mean.SC_1CO, 
                            data=mouse.avs_4MM[(mouse.avs_4MM$sex == "male") & (mouse.avs_4MM$subsp == "Musc"),]

                         , family=binomial(link="logit"))
summary(log.reg_SC.length_mouse.av_1CO)

#2CO
log.reg_SC.length_mouse.av_2CO <- glm(Rec.group ~ mean.SC_2CO, 
                            data=mouse.avs_4MM[(mouse.avs_4MM$sex == "male") & (mouse.avs_4MM$subsp == "Musc"),]

                         , family=binomial(link="logit"))

summary(log.reg_SC.length_mouse.av_2CO)

#3CO
log.reg_SC.length_mouse.av_3CO <- glm(Rec.group ~ mean.SC_3CO, 
                            data=mouse.avs_4MM[(mouse.avs_4MM$sex == "male") & (mouse.avs_4MM$subsp == "Musc"),]

                         , family=binomial(link="logit"))
summary(log.reg_SC.length_mouse.av_3CO)

# bivalent level log regression
log.reg_SC.length_BIV <- glm(Rec.group ~ chromosomeLength, 
                            data=Curated_BivData[(Curated_BivData$subsp == "Musc")& (Curated_BivData$sex == "male"),], family=binomial(link="logit"))

summary(log.reg_SC.length_BIV)

#1CO
log.reg_SC.length_BIV_1CO <- glm(Rec.group ~ chromosomeLength,

                data=Curated_BivData[(Curated_BivData$subsp == "Musc")& (Curated_BivData$sex == "male")&(Curated_BivData$hand.foci.count == 1),], family=binomial(link="logit"))
summary(log.reg_SC.length_BIV_1CO)

#2CO
log.reg_SC.length_BIV_2CO <- glm(Rec.group ~ chromosomeLength, 
                            data=Curated_BivData[(Curated_BivData$subsp == "Musc")& (Curated_BivData$sex == "male")&(Curated_BivData$hand.foci.count == 2),], family=binomial(link="logit"))
summary(log.reg_SC.length_BIV_2CO)

#3CO
log.reg_SC.length_BIV_3CO <- glm(Rec.group ~ chromosomeLength, 
                            data=Curated_BivData[(Curated_BivData$subsp == "Musc")& (Curated_BivData$sex == "male")&(Curated_BivData$hand.foci.count == 3),], family=binomial(link="logit"))

summary(log.reg_SC.length_BIV_3CO)

@@

`r log.reg_SC.length_mouse.av$coefficients`

@@

`r log.reg_SC.length_BIV$coefficients`

#### t.test for SC Length

```{r SC.Length_t.test, echo=FALSE}

T.test_SC.Len_mouse.av <- t.test(mouse.avs_4MM$mean_SC[mouse.avs_4MM$Rec.group == 1],

  mouse.avs_4MM$mean_SC[(mouse.avs_4MM$subsp == "Musc")&(mouse.avs_4MM$sex == "male")&(mouse.avs_4MM$Rec.group == 0)]
)#p-value = 0.001567

#1CO
T.test_SC.Len_mouse.av_1CO <- t.test(mouse.avs_4MM$mean.SC_1CO[mouse.avs_4MM$Rec.group == 1],

  mouse.avs_4MM$mean.SC_1CO[(mouse.avs_4MM$subsp == "Musc") & (mouse.avs_4MM$sex == "male") & (mouse.avs_4MM$Rec.group == 0)]

#2CO
T.test_SC.Len_mouse.av_2CO <- t.test(mouse.avs_4MM$mean.SC_2CO[mouse.avs_4MM$Rec.group == 1],

  mouse.avs_4MM$mean.SC_2CO[(mouse.avs_4MM$subsp == "Musc") & (mouse.avs_4MM$sex == "male") & (mouse.avs_4MM$Rec.group == 0)]

#3CO
T.test_SC.Len_mouse.av_3CO <- t.test(mouse.avs_4MM$mean.SC_3CO[mouse.avs_4MM$Rec.group == 1],

  mouse.avs_4MM$mean.SC_3CO[(mouse.avs_4MM$subsp == "Musc") & (mouse.avs_4MM$sex == "male") & (mouse.avs_4MM$Rec.group == 0)]

#Single Biv level
T.test_SC.Len_single.biv <- t.test(Curated_BivData$chromosomeLength[Curated_BivData$Rec.group == 1],
  Curated_BivData$chromosomeLength[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male")&(Curated_BivData$Rec.group == 0)] )

#1CO
T.test_SC.Len_single.biv_1CO <- t.test(
  Curated_BivData$chromosomeLength[( Curated_BivData$Rec.group == 1) & (Curated_BivData$hand.foci.count == 1) ],

      Curated_BivData$chromosomeLength[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male")&(Curated_BivData$Rec.group == 0) & (Curated_BivData$hand.foci.count == 1) ] 
  )

#2CO
T.test_SC.Len_single.biv_2CO <- t.test(Curated_BivData$chromosomeLength[(Curated_BivData$Rec.group == 1) & (Curated_BivData$hand.foci.count == 2)],

                                       Curated_BivData$chromosomeLength[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male")&(Curated_BivData$Rec.group == 0)&(Curated_BivData$hand.foci.count == 2) ] 
  )

  #3CO
  T.test_SC.Len_single.biv_3CO <- t.test(Curated_BivData$chromosomeLength[(Curated_BivData$Rec.group == 1)&(Curated_BivData$hand.foci.count == 3) ],

  Curated_BivData$chromosomeLength[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male")&(Curated_BivData$Rec.group == 0)&(Curated_BivData$hand.foci.count == 3) ]
  )
```

The mouse average and single bivalents are significantly longer in the high rec group p = `r T.test_SC.Len_mouse.av$p.value` and p=`r T.test_SC.Len_single.biv$p.value`.

@@

#isolate data
#mouse.avs_4MM_male <- mouse.avs_4MM[mouse.avs_4MM$sex == "male",]
#mouse.avs_4MM_M.male <- mouse.avs_4MM_male[mouse.avs_4MM_male$subsp == "Musc",]

@@

rand_SC.Length_MM.male.poly <- exactRLRT(lmer_MM_SC_length_male.poly)#this won't work either...hmm

@@

```{r show.IFD_99, echo=FALSE, warning=FALSE}

@@

```{r IFD.poly_log.reg, echo=FALSE, eval=FALSE}
#This might be a duplicate code chunk
#test. for the high vs low strains and IFD
#mouse.avs_4MM_male

@@

#Single Biv Level
single.biv.1CO_pos_t.test <- t.test(

  Curated_BivData$PER_Foci_1[(Curated_BivData$subsp=="Musc")&(Curated_BivData$sex=="male")&(Curated_BivData$hand.foci.count==1)&(Curated_BivData$Rec.group==1)],

   Curated_BivData$PER_Foci_1[(Curated_BivData$subsp=="Musc")&(Curated_BivData$sex=="male")&(Curated_BivData$hand.foci.count==1)&(Curated_BivData$Rec.group==0)]
  )

@@

#1CO
t.test_mouse.av_siscoten_1CO <- t.test( 
  mouse.avs_4MM$mean.siscoten_1CO[(mouse.avs_4MM$subsp == "Musc")&(mouse.avs_4MM$sex == "male")& (mouse.avs_4MM$Rec.group == 1)],

   mouse.avs_4MM$mean.siscoten_1CO[(mouse.avs_4MM$subsp == "Musc")&(mouse.avs_4MM$sex == "male")& (mouse.avs_4MM$Rec.group == 0)]
  )

#2CO
t.test_mouse.av_siscoten_2CO <- t.test( 
  mouse.avs_4MM$mean.siscoten_2CO[(mouse.avs_4MM$subsp == "Musc")&(mouse.avs_4MM$sex == "male")& (mouse.avs_4MM$Rec.group == 1)],

   mouse.avs_4MM$mean.siscoten_2CO[(mouse.avs_4MM$subsp == "Musc")&(mouse.avs_4MM$sex == "male")& (mouse.avs_4MM$Rec.group == 0)]
  )

#3CO
t.test_mouse.av_siscoten_3CO <- t.test( 
  mouse.avs_4MM$mean.siscoten_3CO[(mouse.avs_4MM$subsp == "Musc")&(mouse.avs_4MM$sex == "male")& (mouse.avs_4MM$Rec.group == 1)],

   mouse.avs_4MM$mean.siscoten_3CO[(mouse.avs_4MM$subsp == "Musc")&(mouse.avs_4MM$sex == "male")& (mouse.avs_4MM$Rec.group == 0)]
  )

####
#Single Biv Level
t.test_single.biv_siscoten <- t.test(

@@

   Curated_BivData$SisCoTen[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male")& (Curated_BivData$Rec.group == 0)]
  )

#1CO
mouse.avs_4MM$mean.siscoten_2CO

#2CO

#3CO
